---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rintchron

<!-- badges: start -->
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![CRAN status](https://www.r-pkg.org/badges/version/rintchron)](https://CRAN.R-project.org/package=rintchron)
[![R build status](https://github.com/joeroe/rintchron/workflows/R-CMD-check/badge.svg)](https://github.com/joeroe/rintchron/actions)
<!-- badges: end -->

rintchron provides an R interface to [IntChron](https://intchron.org)), an indexing service and exchange format for chronological data such as radiocarbon dates ([Bronk Ramsey et al. 2019](https://doi.org/10.1017/RDC.2019.21)).
It includes a user-friendly interface for querying the IntChron database, as well as low level functions for interacting  directly with the IntChron API and other resources using [its schema](https://intchron.org/schema).

## Installation

rintchron has not yet been released on CRAN.
You can install the development version from GitHub using the [remotes](https://remotes.r-lib.org/) package: 

``` r
# install.package("remotes")
remotes::install_github("joeroe/rintchron")
```

## Usage

Use `intchron()` to query databases indexed by IntChron, optionally filtering the results by country and/or site.
For example, to download records from Jordan in the ORAU (`oxa`) and NERC-RF (`nrcf`) databases:

```{r intchron-basic}
library("rintchron")
jordan <- intchron(c("oxa", "nrcf"), countries = "Jordan")
head(jordan)
```

A list of available databases can be retrieved with `intchron_hosts()`:

```{r intchron-hosts}
intchron_hosts()
```

For advanced uses, rintchron also provides a number of functions for interacting with the IntChron API directly; see `vignette("intchron_api")`.

