---
title: "IntChron API"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{IntChron API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library("rintchron")
```

This vignette introduces low-level functions for interacting with IntChron's API directly.
In most cases, the high-level interface provided by `intchron()` should be sufficient; see `vignette("intchron")`.

## Constructing requests to IntChron

## Crawling the IntChron record tree

IntChron's records are organised into a hierarchical tree by data source ('host'), country, and individual data series (typically a site).
For example, the canonical URLs for records from the site of Kharaneh IV (<`r intchron_url("record", "oxa", "Jordan", "Kharaneh IV")`> and <`r intchron_url("record", "nrcf", "Jordan", "Kharaneh IV")`>) encode their position in the tree:

```{r khiv-tree-plot, echo=FALSE}
library("tidygraph", warn.conflicts = FALSE)
library("ggraph", warn.conflicts = FALSE)

tbl_graph(
  nodes = data.frame(name = c("record", "oxa", "nrcf", "Jordan", "Jordan",
                              "Kharaneh IV", "Kharaneh IV"),
                     type = c(rep("index", 5), rep("data", 2))),
  edges = data.frame(from = c(1, 1, 2, 3, 4, 5),
                     to   = c(2, 3, 4, 5, 6, 7))
) %>% 
  ggraph(layout = "tree") +
  geom_edge_link() +
  geom_node_label(aes(label = name, colour = type), label.r = unit(0, "mm")) +
  theme_graph()
```

Here we can distinguish between 'data' records and 'index' records.
Data records include a data series (such as a list of radiocarbon dates) and are found at the terminal nodes of the tree.
Index records are the intermediate nodes in the tree and do not contain a data series, only a list of links to other records.
For example, the index record at `record/oxa/Jordan` (<`r intchron_url("record", "oxa", "Jordan")`>) has a list of all the sites from Jordan in the ORAU database.

There are other index records, such as `host/oxa` and `record/Jordan`, but these are aliases that eventually redirect to the canonical hierarchy described above.

```{r khiv-tree-alias-plot, echo=FALSE}
```

What this means is that if you want to aggregate multiple records, whether a specific subset (e.g. records from Kharaneh IV, records from Jordan), or the whole database, you cannot rely on a single request.
Instead, you need to 'crawl' the tree from an index node to the data nodes you're interested in.

The `intchron_crawl()` function provides this functionality.
Starting from a given index record, it will recursively follow all the links it finds until it reaches the ends of the tree, and return all the data records it finds there.
For example, you can retrieve the entire IntChron database by crawling the whole tree from the `record` entry point:

```{r crawl-everything, eval=FALSE}
# Takes a long time!
everything <- intchron_crawl(intchron_url("record"))
```

The remaining arguments to `intchron_crawl()` give options for refining this search with two common objectives in mind:

1. Crawling the tree to retrieve a whole subset, but with certain types of records excluded.
2. Crawling the tree to find records matching a specific search pattern.

<!--
### Crawling the tree to retrieve a subset


### Crawling the tree to look for specific records

Consider the example of Kharaneh IV above.
The records from this site are found in two databases (ORAU and NERC-RF) and are therefore split across distant branches of the IntChron tree.
To retrieve all the records from this site using the method above, you would have to crawl almost the entire tree and filter the results afterwards, resulting in many unnecessary requests.
It is more efficient to tell `intchron_crawl()` to search for these specific records with ...

With these parameters the crawler only follows links matching the search pattern, greatly reducing the number of requests needed.

## Working with responses
-->
